Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-07-22 paper/paper.tex
Looking at the linear model, but it seems to perform almost as well as the exp-exp model, which is odd

To enable a simulation it is necessary to make the assumptions (in addition to
those implicit in the description of the model above) that i) All
receptor-ligand interactions are repulsive, as for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
all tectal ligand expressions vary exponentially along a given axis, following
Eq.\,\ref{e:tecexp}. The result of running this system are presented in
Fig.\,\ref{f:wt}.
%
Fig.\,\ref{f:wt}Ai shows a dual-colour map of retinal ganglion cell positions
on the square retina, with red encoding position on the $x$ axis (temporal to
nasal) and green encoding position on the $y$ axis (ventral to
dorsal). Fig.\,\ref{f:wt}Aii presents a fishnet plot of the expected location on
the tectal surface of each RGC at the end of the development process. Thus,
the `red' RGC from the ventro-nasal corner should map to the caudo-medial
corner of the tectum and the `green' RGC from the dorso-temporal corner should
find its way to the rostro-lateral corner on the tectum. The lines between the
dots of Fig.\,\ref{f:wt}Aii indicate the adjacency relationships between
cells. Thus, colour in each panel indicates a branch or axon's location of
origin on the retina.
%
Fig.\,\ref{f:wt}B shows the initial state of the system, with branches
stochastically arranged around the rostral edge of the tectum. Each dot
represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}C and \ref{f:wt}D
show the development of the system with time, with Fig.\,\ref{f:wt}D showing a
well ordered arrangement, with a majority of adjacency relationships matching
the experimental prediction (i.e. there are few crossed lines). The
intermediate timepoint at $t=40$ indicates that branches quickly cover the simulated
tectal surface, with final adjacency relationships developing over a longer
timescale. This dynamic behaviour is also evident in Fig.\,\ref{f:wt}E, which
shows \textbf{RMS error or some metric} of the axon centroids with respect to
the experimental prediction. Fig.\,\ref{f:wt}F shows the final
location of each branch in the simulation, which gives a more continuous view
of the order present in Fig.\,\ref{f:wt}D. Finally, Fig.\,\ref{f:wt}G shows the
branches and their centroid path history for five selected RGCs, with path
histories strongly resembling those presented by \citet{simpson_simple_2011}.

@@

\begin{figure}
\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
\caption{Surgical manipulations}
\label{f:manip}
\end{figure}

>>>>

To enable a simulation it is necessary to make the assumptions (in addition to
those implicit in the description of the model above) that i) All
receptor-ligand signalled interactions are repulsive, as for EphA-ephrinA
coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
all tectal ligand expressions vary exponentially along a given axis, following
Eq.\,\ref{e:tecexp}. The result of running this system are presented in
Fig.\,\ref{f:wt}.
%
Fig.\,\ref{f:wt}Ai shows a dual-colour map of retinal ganglion cell positions
on the square retina, with red encoding position on the $x$ axis (temporal to
nasal) and green encoding position on the $y$ axis (ventral to
dorsal). Fig.\,\ref{f:wt}Aii presents a fishnet plot of the expected location on
the tectal surface of each RGC at the end of the development process. Thus,
the `red' RGC from the ventro-nasal corner should map to the caudo-medial
corner of the tectum and the `green' RGC from the dorso-temporal corner should
find its way to the rostro-lateral corner on the tectum. The lines between the
dots of Fig.\,\ref{f:wt}Aii indicate the adjacency relationships between
cells. Thus, colour in each panel indicates a branch or axon's location of
origin on the retina.
%
Fig.\,\ref{f:wt}B shows the initial state of the system, with branches
stochastically arranged around the rostral edge of the tectum. Each dot
represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}C and \ref{f:wt}D
show the development of the system with time, with Fig.\,\ref{f:wt}D showing a
well ordered arrangement, with a majority of adjacency relationships matching
the experimental prediction (i.e. there are few crossed lines). The
intermediate timepoint at $t=40$ indicates that branches quickly cover the simulated
tectal surface, with final adjacency relationships developing over a longer
timescale. This dynamic behaviour is also evident in Fig.\,\ref{f:wt}E, which
shows \textbf{RMS error or some metric} of the axon centroids with respect to
the experimental prediction. Fig.\,\ref{f:wt}F shows the final
location of each branch in the simulation, which gives a more continuous view
of the order present in Fig.\,\ref{f:wt}D. Finally, Fig.\,\ref{f:wt}G shows the
branches and their centroid path history for five selected RGCs, with path
histories strongly resembling those presented by \citet{simpson_simple_2011}.

@@

Fig.\,\ref{f:manip} shows the result of simulating three surgical
manipulations alongside the wildtype result, which is shown on Row A. Rows B
and C show graft rotation simulations (90$\degree$ and 180$\degree$). In each
case an 8x8 square on the tectum (each unit is 0.05x0.05) is rotated. The
rotational sense can be observed clearly in the `Branches' column; for the
90$\degree$ rotation, the top-right (caudal-lateral) corner of the rotated
square is red, whereas the top-left corner of the wildtype tectum is red. \textbf{etc}.

\begin{figure}
\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
\caption{Surgical manipulations. Row A: wildtype. Row B: 90 degree
rotation. Row C: 180 degree rotation. Row D: Graft swap.}
\label{f:manip}
\end{figure}

\subsection*{ligand expression profiles}

A section to show the effect of assuming linear or logarithmic expression
profiles. Especially that linear expression fails to reproduce the rotation
and graft swap manipulations. Hmm, actually, it looks ok?!

\subsection*{Attractive receptor-ligand interactions}

Show that, if you switch one of the receptor-ligand  interactions in the
system to be attractive instead of repulsive, then you have to make the ligand
expression form logarithmic instead of exponential.
